---
title: "ProblemSet2"
author: "Aaron Graybill"
date: "`r Sys.Date()`"
output: 
  html_document: 
    toc: yes
  pdf_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(dplyr)
library(tidyr)
library(ggplot2)

source(here("R-scripts","ggplotTheme copy.R"))
```

## Problem 2.1

```{r 2.1 data pull}
df <- 
  read.csv(here('Data',"playbill.csv"))
movie_lm <- 
  lm(CurrentWeek~LastWeek,data=df)
summary(movie_lm)
```

### 2.1.a
We can find the 95% confidence interval on the $\hat{\beta}_1$ term with the following code:
```{r 2.1.a}
confint_2.1 <- 
  confint(movie_lm)
```

The confidence interval around $\hat{\beta}_1$ is $(`r confint_2.1[2,]`)$ which actually contains $1$, so we do not have significant evidence that the true $\beta_1\neq1$. We can compute this test a different way. We can run a hypothesis test where $H_0\colon \beta_1=1$ and $H_a\colon \beta_1\neq1$. We showed in class that:
$$
\frac{\hat{\beta}_1-1}{se(\hat{\beta}_1)}\sim t_{n-2}
$$
We have all of these values from our regression output, so we can compute the $p$-value that the null hypothesis is true in the following way:

```{r 2.1.b 1}
test_stat=(movie_lm[["coefficients"]][["LastWeek"]]-1)/coef(summary(movie_lm))[, "Std. Error"]["LastWeek"]

critical_region <- 
  qt(c(.025,.975),movie_lm$df.residual)

p_val <- (pt(test_stat,movie_lm$df.residual)*2)
  
```

The 95% critical region, as computed is $(-\infty,-2.12)\cup(2.12,\infty)$, and our test statistic is equal to $`r test_stat`$ is not in that region, so we have insufficient evidence to reject the null hypothesis that $\beta_1=1$ at the 95% level. In fact, the $p$-value from this test is $`r p_val`$ which is well above the required $.05$ at the 95% level.

### 2.1.b

As proven in class:
$$
\frac{\hat{\beta}_0-\beta_0^0}{se(\hat{\beta}_0)}\sim t_{n-2}
$$
In this case our $H_0$ is $\beta_0=1000$ and $H_a$ is $\beta_0\neq1000$. Implementing similar code to above gives:

```{r 2.1.b 2}
test_stat=(movie_lm[["coefficients"]][["(Intercept)"]]-1000)/coef(summary(movie_lm))[, "Std. Error"]["(Intercept)"]

critical_region <- 
  qt(c(.025,.975),movie_lm$df.residual)

p_val <- (pt(-abs(test_stat),movie_lm$df.residual)*2)
```

I have to do some business with `-abs(test_stat)` to ensure that when I compute the cumulative density it's in the left tail so the $p$-value is just two times the computed density. Anyway, computing the $p$-value gives: `r p_val` which is quite high and provides very little evidence that the true $\beta_0\neq1000$. In fact since $p>.5$, there is more evidence that $H_0$ is true than the alternative.

### 2.1.c
We showed in class that the prediction interval is given by:
$$
\hat{\beta}_0+\hat{\beta}_1x^*\pm t_{\frac{\alpha}{2},n-2}S\sqrt{\frac{1}{n}+\frac{x^*-\overline{x}}{SXX}}
$$
But thankfully `R`will take care of that for us with the following code:
```{r 2.1.c}
prediction_points <- 
  data.frame(LastWeek=400000)
prediction <- 
  predict(movie_lm,prediction_points,interval = 'prediction',level=.95,se.fit=T)
prediction
```
Summarizing those results, the point estimate for `CurrentWeek` is $\$`r prediction[['fit']][1] %>% format(scientific=F)`$ with a 95% prediction interval of: $(`r prediction[['fit']][2] %>% format(scientific=F)`,`r prediction[['fit']][3] %>% format(scientific=F)`)$. So we are 95% percent certain the true value of $y^*$ would lie in the aforementioned range.

## Problem 2.2

